% vim: ts=2:sw=2:tw=80:et
\thispagestyle{fancy}
\pagestyle{fancy}

\section{Prerequisites\index{Installing prerequisites}}

This project depends on several other python-related projects and these must
first be installed.

\subsection{Dependencies}
Arbwave is dependent on various packages, both generally available and also
several that are distributed with the arbwave project repository.  This list
enumerates the various specific packages that Arbwave depends on.  Those
distributed as a part of the Arbwave project repository are marked below.

\begin{itemize}
  \item Python packages/libraries:
    \begin{itemize}
      \item \textbf{Python ($\geq$3.4)}
      \item \textbf{python3-gi} \\
        \textbf{gi} is the base package of the GObject instropsection library
        interface.  This library provides dynamic access to various other
        libraries, including the Gtk runtime.  In addition to installing
        \textbf{gi}, one must also have these installed (often already installed
        on Gnome-based Linux systems):
        \begin{itemize}
          \item Gtk $> 3.10$
          \item Pango $> 3.10$
        \end{itemize}

      \item \textbf{matplotlib} \\
        This package also provides mpl\_toolkits import.
      \item \textbf{numpy} \\
        Generally, this is included as part of scipy and required for matplotlib.
      \item \textbf{scipy} \\
        scipy.interpolate imports are required for channel scaling/calibration.
      \item A graph library for graph-analysis of connections between hardware
        components.\\
        You must have either of these installed:
        \begin{itemize}
          \item \textbf{python3-igraph} (\underline{Significantly} faster)
          \item \textbf{networkx} (Apparently more available)
        \end{itemize}

      \item \textbf{Pyro4}

      \item \textbf{pylibnidaqmx}
        \begin{itemize}
          \item must have modifications by S. Olson
          \item Python wrapper library (wraps NI DAQmx nicaiu.dll c library)
            provides actual linkage to the NIDAQmx drivers
          \item Only needed if NIDAQmx relevant hardware is going to be used
        \end{itemize}

      \item \textbf{python3-comedilib}
        \begin{itemize}
          \item Only for Linux installs.
          \item comedi drivers in Linux kernel staging tree must be installed.
          \item Must use the new CTypes interface (not the swig interface).
        \end{itemize}

      \item \textbf{physical}
        \begin{itemize}
          \item http://github.com/olsonse/physical
          \item dimensional analysis
          \item units and constants library (c++, python, octave, gnuplot)
        \end{itemize}

      \item \textbf{viewpoint}
        \begin{itemize}
          \item python wrapper library (wraps Viewpoint dio64\_32.dll c library)
          \item only required if using ViewpointUSA dio64 hardware
        \end{itemize}

      \item \textbf{marvin}
        \begin{itemize}
          \item python wrapper library around MarvinTest c-libraries
          \item only required if using MarvinTest FPGA or analog output hardware
        \end{itemize}
    \end{itemize}

  \item C libraries:
    \begin{itemize}
    \item \textbf{NI DAQmx (nicaiu.dll)} [Windows] \\
      Only needed if NIDAQmx relevant hardware is going to be used.
    \item \textbf{Viewpoint DIO64 (dio64\_32.dll)} [Windows] \\
      Only needed if Viewpoint's DIO64 digital output/input board will be used.
    \item \textbf{Marvin GxFpga (libGxFpga.so, or GxFpga.dll)} [Windows] \\
      Only needed if MarvinTest GxFpga with \acro{AFRL} Timing-Generator
      firmware is used.
    \item \textbf{Marvin GxAo (libGxAo.so, or GxAo.dll)} [Windows] \\
      Only needed if MarvinTest GxAo analog output hardware is used.
    \item \textbf{MarvinTest GtLinux} [Linux] \\
      This package of libraries and kernel module support all relevant Marvin
      test hardware.
    \item \textbf{comedilib} [Linux] \\
      Supports many hardware pieces from various manufacturers under Linux.
      Currently, only NI hardware is known to have the requisite interfaces
      implemented to use with Arbwave.
    \end{itemize}
\end{itemize}

\section{Installation\index{Installing Arbwave}}
Since Arbwave is a set of Python scripts and other related code that has certain 
dependencies and requirements,
it is not feasible to write a strict set of instructions for ``installing'' Arbwave,
but the following is a set of guidelines that might be used to make Arbwave operational.

\subsection{Windows Platforms}

Note that you may need to specify a web proxy;\\
the commands to get conda to use a web proxy, in some cases, is:
\begin{verbatim}
set http_proxy=http://proxy.server:port
set https_proxy=http://proxy.server:port
\end{verbatim}


The following are a set of steps used to make Arbwave functional on a Windows system.

\begin{enumerate}
  \item Install MSYS2

  \item Open a MSYS2 shell and install Python 3 \\
    MSYS2 uses pacman as package manager.  When using a 64-bit Windows the
    command is
    \begin{verbatim}
    pacman -S mingw-w64-x86_64-python3
    \end{verbatim}

  \item Install GTK+ 3 \\
    This step is following instructions provided by GTK at
    http://www.gtk.org/download/windows.php
    \begin{itemize}
      \item Install GTK3 and its dependencies:
      \begin{verbatim}
      pacman -S mingw-w64-x86_64-gtk3
      \end{verbatim}

      \item Install GTK core applications:
      \begin{verbatim}
      pacman -S mingw-w64-x86_64-devhelp
      \end{verbatim}

      \item Install the Python bindings:
      \begin{verbatim}
      pacman -S mingw-w64-x86_64-python3-gobject
      \end{verbatim}

      \item Install build tools:
      \begin{verbatim}
      pacman -S mingw-w64-x86_64-toolchain base-devel
      \end{verbatim}
    \end{itemize}

  \item Install necessary Python packages/libraries (scipy, numpy, Pyro4): \\
  Scipy is available as MSYS2 package, but Pyro4 not. Therefore package
  manager \bf{pip} was also installed:
  \begin{verbatim}
  pacman -S mingw-w64-x86_64-python-scipy
  \end{verbatim}
  \begin{verbatim}
  pacman -S mingw-w64-x86_64-python-pip
  \end{verbatim}

  The other packages were installed using pip:
  \begin{verbatim}
  pip install Pyro4
  \end{verbatim}

  \item Install graph library \\
  It seems that a clean Windows install of \textbf{python3-igraph}
  was not trivial.  However, note that taking the time to install
  \textbf{python3-igraph} will result in some performance improvement, though
  not necessarily noticable.

  Arbwave (1.1.0 as of 2020.01.30) is not compatible with networkx 2.x, so
  version 1.11 has to be installed:
  \begin{verbatim}
  pip install python-networkx==1.11
  \end{verbatim}

  \item Do either of these:
    \begin{enumerate}
      \item  (start cmd.exe)
        \begin{enumerate}
          \item cd path/to/arbwave-project
          \item Execute 'python3 project-shell'
          \item cd arbwave/
          \item 'python3 run.py [options]'\\
            use \verb|--help| to see options
        \end{enumerate}
      \item
        \begin{enumerate}
          \item Put all the other packages in your PYTHONPATH somehow
          \item cd path/to/arbwave-project/arbwave/
          \item 'python run.py [options]'\\
            use \verb|--help| to see options
        \end{enumerate}
    \end{enumerate}
\end{enumerate}



\subsection{Ubuntu}
Ubuntu makes this very easy.  To install most of the non-project components, one
simply does:

\begin{verbatim}
apt-get install \
  python3-matplotlib python3-scipy python3-igraph python3-pyro4 \
  python3-gi libgtk-3-dev libpango1.0-dev
\end{verbatim}


Then, very similar to the last part of the Windows guidelines: \\
Do either of:
\begin{enumerate}
  \item
    \begin{enumerate}
      \item cd path/to/arbwave-project
      \item Execute project-shell
      \item cd arbwave/
      \item ./run.py [options]\\
        use \verb|--help| to see options
    \end{enumerate}
  \item
    \begin{enumerate}
      \item Put all the other packages in your PYTHONPATH somehow
      \item cd path/to/arbwave-project/arbwave/
      \item ./run.py [options]\\
        use \verb|--help| to see options
    \end{enumerate}
\end{enumerate}
